---
title: "A/B testing, Bayes style"
output:  html_document
---

```{r, echo=FALSE}
source('bayesianABTesting.R')
```

###But why?
So that you understand your test results correctly, including the uncertainty in them, without the p-value hassle.

###Also:

Gentle introduction to Bayesian methods.

###After this
You understand the results of Bayesin analysis, and are familiar with terms such as _posterior_,  _prior_, and _statistical model_.

***

###Our case

Reaktor website has a one-click purchase of a fixed-price-scope schedule-team agile software project, or whatever. More to the point, there are _two variants of the button_. Which is better? How much?

**What do we measure**? 

Visits on the page and button clicks.

Below are some simulated data, with pre-set click rates. Those are the underlying rates that we do not know in the measurement situation, and which we are after. Per day, we have on average 1500 visits on the page, of which 20% see the option __B__, for a total of six days.

Note that if you re-knit this document from the R Markdown file, data is re-generated and the results will change. How clear the difference between __A__ and __B__ looks like, from the raw click rates, _varies by run_. _That's also how it is in real life!_ Our goal is to get beyond that variation to get an idea of how things are in reality. 

Reality here refers to the pre-set click rates. In practice we also have pre-set click rates, but they are set by  nature. We do not observe them directly. We observe only data.

```{r}
rate.A <- 0.052
rate.B <- 0.068

example <- click.data(rate.A, rate.B, mean.visits=1500, test.group=0.2, days=6)

print(example)
```

So this is the (simulated) setup, and the data above is all we know from measurements. 

---

That's how the data looks like as plots:

```{r}
example.plots <- click.plots(example)
print(example.plots$histograms)
print(example.plots$rates)
```

---
###Let's mathematics

What do we want to know?

The _probabilities_ __P~A~__ and __P~B~__that the visitor buys a project when shown a button __A__ or __B__. In this case they are equal to click rates __r~A~__ and __r~B~__ 

What do we have?

Click (__k__) and visit (__N__) counts

---

###A statistical model

Think how the data is generated.

If the probability to click is __P__ the probability to get __k__ clicks in __N__ visits is (the binomial distribution)

$$ P(k,N) = Bin(n,K) \sim P^k (1-P)^{N-k} $$

This gives me way to calculate expected number of clicks __k__ in __N__ visits when I know __P__.

Sadly, I want to have this the other way round: I want to know the expected probability of __P__, when I know that there are __k__ clicks in __N__ visits.

---

###Enter Bayes

The way to turn probabilities around.

$$Prob(k,N,P) = Prob(k, N | P) Prob(P) = Prob(P | k, N) Prob(k, N) $$
$$\Rightarrow Prob(P | k, N) = \frac{Prob(k, N | P) Prob(P)}{Prob(k, N)} $$

Essentials: 

I can turn the probabilities of getting __k__ clicks in __N__ visits when I know the click rate __r__, into probability of click rate being __r__ when I know that there was __k__ clicks in __N__ visits. 

There is one more thing: The probability of click rate being __r__ in the first place. This is my prejudice or domain knowledge of the click rate. If I know nothing, I'll set it to constant, meaning: all values are equally good guesses. (We'll come back to this later)


```{r}
example.posterior <- abtest.posteriorplot(example)
# quartz(width=14, height=6)
print(example.posterior$plot
      + geom_vline(xintercept=rate.A, linetype='dashed', colour=colours[1])
      + geom_vline(xintercept=rate.B, linetype='dashed', colour=colours[2]))

```


Matemaattisesti: mik?? on todenn??k??isyys p ostaa varianteilla
Oikea kysymys : mik?? on p?
Malli: Binomiaali jakauma
kuinka monta konversiota N n??yt??ll???
Malli sille miten ajatellaan datan generoituneen
v????rinp??in (data ehdolla malli)
Bayesin kaava (johto, saa olla ym??rt??m??tt??)
Priori vakioksi, palataan my??h
Posteriori plotteja -> kaikki ymm??rt???? mik?? on posteriori. 
todenn??k??isyysjakaumat mitta meid??n ep??varmuudelle
Posteriorista voidaan laskea hakutut asiat
kuinka paljon toinen on parempi (keskinm????rin), kuinka varmaa on ett?? on parempi
Priori
Katotaa Bayesin kaavaa
prioriin pakko laittaa jotain
priori on AINA, my??s frekventisteill??, vakkei ne aina my??nn??
Meid??n ennakkoluulo asiasta
Luonnollista laittaa tasainen -> ei tiedet?? mit????n
toinen tapa, laitetaan matemaattisesti kivan muotoinen funktio -> funktio ei muuta muotoa kun lis??t????n likelihood
konjugaatti priori
t??ss?? Beta
Esimerkki, tehd????n sama analyysi Beta priorilla
Eri countit nyt
Poll: mik?? on konversio todenn??k??isyys?
Poll: paljonko haluat dataa ett?? hylk????t ennakkoluulon?
Betan parametrit tulee t??st?? (selitet????nk???)
Betan keskiarvo ja hajonta (Poll: onko informatiivista?)
Uusia posterioiri plotteja

Huomaa:
Laskut n??in helppoja vain t??ss?? tapauksessa, hyvin harvoin -> Syy miksi Bayes ei ole yleistynyt
Nyky????n tietskarit tekee simulaatioilla, ei en???? niin vaikeeta

Extroja:
Frekventistisi?? tuloksia, milt?? n??ytt????? N??m??kin on oikein, eri tapa esitt????
Monta ryhm???? (A/B ei viittaa ryhmien m????r????n vaan outcomien m????r????n) monta treatment
kovariaatit -> probabilistig programming, ratkaisee ongelman todenn??k??isyyden k????n??st?? 
Testiryhmien tekeminen ja kontrollointi T??RKE????.
Menee helposti pieleen.
Tapauskohtaista.
ajalliset vaikutukset
