---
title: "A/B testing, Bayes style"
output:
  ioslides_presentation:
    widescreen: no
---

```{r, echo=FALSE, message=F}
source('bayesianABTesting.R')
```

## But why?
So that you understand your test results correctly, including the uncertainty in them, without the p-value hassle.

### Also:

Gentle introduction to Bayesian methods.

### After this
You understand the results of Bayesin analysis, and are familiar with terms such as _posterior_,  _prior_, and _statistical model_.

## Our case

Reaktor website has a one-click purchase of a fixed-price-scope schedule-team agile software project, or whatever. There are _two variants of the button_. 

Which is better? How much?

**What do we measure**? 

Visits on the page and button clicks.

**What do we want to know**?

The probablity that the visitor clicks on the button to buy

= click rate 

= conversion rate


## Collected data

Simulated data for a day, average 2000 visit on the page, 20% see option _B_
<div class="centered">
```{r, echo=FALSE}
rate.A <- 0.049
rate.B <- 0.068

example <- click.data(rate.A, rate.B, mean.visits=2000, test.group=0.2, days=1)

print(example)
```
</div>
## Bayesian result

<div class="centered">
Posterior probability distribuition of click rates
```{r, echo=FALSE}
example.posterior <- abtest.posteriorplot(example)

print(example.posterior$plot + xlab('click rate') + ylab('') 
      + theme_minimal())
```
</div>

## Problems with the usual way

I get:

- Average conversion rate (_ok, I get this_)
- P-value (_probablity of obtaining the data when null hypothesis is true ... WTF_)
- Z-score (_no idea what this is_)
- Confidence intervals (_I've been told that nobody understands_)

Loads of other numbers I don't understand

## What's behind Bayesian result

Assumption: visitor on page has probablity __P~A~__ of clicking the button when shown variant __A__ and __P~B~__ when shown __B__. Visits are independent of each other

- For those interested, this is called binomial model

Before collecting any data I have an opinnion what the click rate is. Here the opinnion was: I know nothing, all values (between 0 and 1) __P~A~__ and  __P~B~__ are equally good.

- The opinnion _must_ be included in the analysis, even when it is "I don't know"
- This is called the _prior_

## What do the results mean

Exactly what you think they do (maybe).

The meaning of the of click rate distribution is exactly the same thing as when you say that your _estimate_ of the rain probability for tomorrow is between 20% and 30%.

## The cycle

State your knowledge about the click rates. Can be "_I know nothing_". This is your
_prior_ distribution.

Define your model how data is generated in your process (Visitors have a probability __P__ to click the button). </br> Ask your local Scientology dealer for help

Collect data

Do Bayesian analysis (or ask someone to do it for you), it is the method of updating the
knowledge with the collected data 

You get _posterior_ distribution, the knowledge we have about the click rates _given that we have the observed data_. 

## Fun facts

Prior and posterior distributions mean the same thing: it is the mathematical 
representation of our knowledge including the uncertainty in it (narrow or wide distribution).

In the next experiment (or next day of same experiment) we can, and should use the previous 
posterior as our prior

Posterior (or prior) contain all the information we have and everything can be calculated from it. 
Like "How much __A__ is better than __B__?", "How certain we are that __A__ is really better 
than __B__" etc.

## From clicks to posterior

<div class="centered">
```{r, echo=FALSE}
rate.A <- 0.049
rate.B <- 0.068

example <- click.data(rate.A, rate.B, mean.visits=1500, test.group=0.2, days=6)

print(example)
```
</div>

---

<div class="centered">
```{r, echo=FALSE}
example.plots <- click.plots(example)
print(example.plots$histograms + theme_minimal())
```
</div>

---

<div class="centered">
```{r, echo=FALSE}
print(example.plots$rates + theme_minimal())
```
</div>

---

<div class="centered">
```{r, echo=FALSE}
example.posterior <- abtest.posteriorplot(example, prior=c(5,95), plot.prior=TRUE)

print(example.posterior$plot + xlab('click rate') + ylab('') 
      + theme_minimal())
```
</div>

